---
title: "SNA Example Cleaning Template- Fellow"
author: "Rebekah Jacob"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float: true
    theme: flatly
    code_folding: show
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

# Introduction

This serves as documentation to the cleaning and management of the social network survey featuring fictional "Fellows" from a research fellowship program.

Libraries
```{r}
library(tidyverse) #Data management
library(DT) #html tables to display data
library(qualtRics) #read Qualtrics data
#library(igraph) making a note here that we'll add more later
```

# Survey data
```{r}
survey<- read_survey("Example survey/Example-SNA-survey-data.csv") %>%
  mutate(full_name=paste0(RecipientLastName, ", ", RecipientFirstName)) %>%
  select(ID, SortID, full_name, position:connect_rating) %>%
  arrange(SortID)

survey
```

# Attributes

We have a few sources of attribute data. One is assigned info from admin data. The other is anything from surveying. A note to remember that surveys will likely have missing data if people didn't complete certain questions and/or if not a 100% response rate. We need to make sure to include SortID. This helps us take out attributes from anyone that wasn't included in that year's survey.

Since this is test data, we have few missing data. In reality, even if you get an 80% response to an SN survey (still shooting high), you'd be missing 20% of your attribute data.
```{r}

qual_panel<- read_csv('Contacts/Qualtrics-contact-list.csv') %>%
  select(ID, SortID, referred, location) #Just need SortId for merging and any variables that are NOT in survey


#from survey data
survey_att<- survey %>%
  select(ID, position, connect_rating, full_name) #this is where you pull out any variables that are NOT network questions (people names)


atts<- list(qual_panel, survey_att) %>%
  reduce(full_join, by="ID") %>%
  filter(!is.na(SortID)) #only keep those with a sort order because those are the only ones in 2020's survey

head(atts)
```

## Recodes for groups

Within the attribute file, we may be interested in grouping some categories ahead of time if we know that we need them for stats and visualizations later.

For example, I'm going to go ahead and collapse the `connect_rating` varible to a binary variable: *"Easy"* vs. *"Difficult"*. 
```{r}
atts<- atts %>%
  mutate(connect_rating_bi=case_when(
    grepl("difficult", connect_rating)~"Difficult",
    grepl("easy", connect_rating)~"Easy",
  ))

atts %>%
  group_by(connect_rating_bi) %>%
  tally()
```

## Adding colors

At this point, we can decide on a few variables that we might want to show on visualizations.

You can find all kinds of fun palettes online. I used [Huemint](https://huemint.com/brand-3/) for this scheme.
```{r}
atts<- atts %>%
  mutate(
    referred_color=case_when(
      referred=="Friend"~"#15d4f7",
      referred=="Boss"~"#f96e26", 
      referred=="Coworker"~"#5e53f7"),
    
    location_color=case_when(
      location=="Mex"~"#15d4f7", 
      location=="US"~"#f96e26",
      location=="Canada"~"#5e53f7"), 
    
    position_color=case_when(
      position=="Fellow"~"#15d4f7", 
      position=="Mentor"~"#f96e26"), 

    connect_rating_bi_color=case_when(
      connect_rating_bi=="Easy"~"#282728",
      connect_rating_bi=="Difficult"~"#f96e26" 
    )
    )
```


## Save attribute file

since we did quite a bit of work on this data, I'm going to save it as an Excel file so we can access it outside of this code. For example, we can easily use for quick descriptives later if we chose.
```{r}
write_csv(atts, "Contacts/attribute-file.csv")
```


# Network data

We've already loaded the survey data above in order to get attribute data, etc. Now let's pull out the network data from the survey.
The survey data have 23 respondents, but our attribute data is for 26.
```{r}
nrow(survey)
nrow(atts)
```

## Adding missing rows

Since not everyone completed the survey (not 100% response), we need to add in rows for those that didn't complete. We have to use the attribute data to do this (which includes ALL IDs).

**We do this BEFORE we split out the separate activity networks**

We're back to having 26 members, which is what we should have if we had a 100% response.
```{r}
survey<- survey %>%
  full_join(atts %>%
               select(ID, SortID, location, referred, connect_rating_bi),
             by=c("ID","SortID")) %>%
  arrange(SortID)

nrow(survey)
```

## Recode to numeric

Here are the basic transformations here:

**Contact**: `contact_1` - `contact_26` (fellows) and Q6 (faculty) contact:

- I don't know this person ~ NA
- I know who this person is but we have had no contact ~ NA
- I have had contact with this person yearly ~ 1
- I have had contact with this person quarterly ~ 2
- I have had contact with this person monthly ~ 3

**Collaboration**:
- Engaged in a joint research project: `collab_1_1` - `collab_26_1`, 
- Published or wrote a scholarly manuscript-`collab_1_2` - `collab_26_2`,
- Engaged in joint grant writing- `collab_1_3` - `collab_26_3`,
- Co-taught a course or training- `collab_1_4` - `collab_26_4`,
- Presented research- `collab_1_5` - `collab_26_5`,
- I did not engage in any of these activities- `collab_1_6` - `collab_26_6`

These will not be weighted. They will be used later to break out separate activity networks. So for now, I will just assign 1 to each nomination across the collaboration variables.

They will also be undirected, meaning we won't worry about the direction of the collaboration (in or out), we will symmetrize the data.


To start, I'm going to name vars of interest in list so that I can use them now and later. Then I can pass the function to the list of vars.

Notice I leave "No contact" out of my function because we will ignore as NA.
```{r}
contact_vars<- survey %>%
  select(starts_with("contact")) %>%
  names()

recode_freq_fun<- function(x){
  case_when(
    x=='I have had contact with this person yearly' ~ 1,
    x=='I have had contact with this person quarterly' ~ 2,
    x=='I have had contact with this person monthly' ~ 3
  )
}



survey<- survey %>%
  mutate_at(contact_vars, recode_freq_fun)

head(survey %>%
  select(contact_vars))
```


```{r}
collab_vars<- survey%>%
  select(starts_with("collab")) %>%
  names()

recode_nom_fun<- function(x){
  as.numeric(case_when(
    x=="Engaged in a joint research project"~1,
    x=="Published or wrote a scholarly manuscript"~1,
    x=="Engaged in joint grant writing"~1,
    x=="Co-taught a course or training"~1,
    x=="Presented research"~1,
    x=="I did not engage in any of these activities"~1, #we won't select this as a network, but just keeping it for now
    x=="They mentored me"~1, #combining into just one function to include mentor nominations
    x=="I mentored them"~1 #combining into just one function to include mentor nominations
  ))
}

survey<- survey %>%
  mutate_at(collab_vars, recode_nom_fun)

head(survey %>%
  select(collab_vars))
```

## Matrices

Now I can separate each df into separate activity network

Contact
```{r}
contact<- survey %>%
  arrange(SortID) %>%
  select(contact_vars, -(c(ID, SortID))) %>%
  as.matrix() %>%
  `colnames<-`(atts$ID) %>%
  `rownames<-`(atts$ID)

```

Research
```{r}
research<- survey %>%
  arrange(SortID) %>%
  select((starts_with("collab") & ends_with("1")), -(c(ID, SortID))) %>%
  as.matrix() %>%
  `colnames<-`(atts$ID) %>%
  `rownames<-`(atts$ID)
```

Manuscript
```{r}
manuscript<- survey %>%
  arrange(SortID) %>%
  select((starts_with("collab") & ends_with("2")), -(c(ID, SortID))) %>%
  as.matrix() %>%
  `colnames<-`(atts$ID) %>%
  `rownames<-`(atts$ID)
```

Grant
```{r}
grant<- survey %>%
  arrange(SortID) %>%
  select((starts_with("collab") & ends_with("3")), -(c(ID, SortID))) %>%
  as.matrix() %>%
  `colnames<-`(atts$ID) %>%
  `rownames<-`(atts$ID)
```

Teach
```{r}
teach<- survey %>%
  arrange(SortID) %>%
  select((starts_with("collab") & ends_with("4")), -(c(ID, SortID))) %>%
  as.matrix() %>%
  `colnames<-`(atts$ID) %>%
  `rownames<-`(atts$ID)
```

Present
```{r}
present<- survey %>%
  arrange(SortID) %>%
  select((starts_with("collab") & ends_with("5")), -(c(ID, SortID))) %>%
  as.matrix() %>%
  `colnames<-`(atts$ID) %>%
  `rownames<-`(atts$ID)
```


## Symmetrize

For nondirected networks (contact and collab activites) we will make connection symmetric (if A connected to B, then B is connected to A)

For the contact network, for tie breakers, we'll take the lower rating. If A says monthly and B says yearly, we go with a yearly connection between A and B.

Then we replace all NA values with 0.

Here's a function to do all.
```{r}
sym_fun<- function(mat){
  tmat<- t(mat) #transpose
  symat<- pmin(mat, tmat, na.rm=TRUE) #parallel minimum function and if NA, then take other person's rating
  symat %>%
  replace(., is.na(.), 0) #replace all NAs with 0's so not missing
}
```

```{r}
contact<- sym_fun(contact)
```

```{r}
research<- sym_fun(research)
```

```{r}
manuscript<- sym_fun(manuscript)
```

```{r}
grant<- sym_fun(grant)
```

```{r}
teach<- sym_fun(teach)
```

```{r}
present<- sym_fun(present)
```

Here I'm creating one network to summarize *any* collaboration activity. We add the matrices and then convert any tie to 1 (otherwise, if folks had multiple connections, it would be larger than 1).
```{r}
anycoll<- research+manuscript+grant+teach+present

anycoll[anycoll > 0] <-1
```

# Convert to network object

Now we can load igraph packages. Also, pay attention to the options for when each is being converted.
```{r}
library(igraph)
```

Contact
```{r}
contactnet<- graph_from_adjacency_matrix(contact,
                                              mode="undirected", #Symmetrized
                                              weighted = TRUE, #Contact has weights, but not other networks in this set
                                              diag = FALSE) #do not want loops

contactnet
```

Checking that node name matches our sortID, they do and they are character format.
```{r}
vertex_attr(contactnet, "name")
```

```{r}
anycollnet<- graph_from_adjacency_matrix(anycoll,
                                              mode="undirected", #Symmetrized
                                              weighted = NULL, #no weights
                                              diag = FALSE) #do not want loops

anycollnet
```


Research
```{r}
researchnet<- graph_from_adjacency_matrix(research,
                                              mode="undirected", #Symmetrized
                                              weighted = NULL, #no weights
                                              diag = FALSE) #do not want loops

researchnet
```

Manuscript
```{r}
manuscriptnet<- graph_from_adjacency_matrix(manuscript,
                                              mode="undirected", #Symmetrized
                                              weighted = NULL, #no weights
                                              diag = FALSE) #do not want loops

manuscriptnet
```
Grant
```{r}
grantnet<- graph_from_adjacency_matrix(grant,
                                              mode="undirected", #Symmetrized
                                              weighted = NULL, #no weights
                                              diag = FALSE) #do not want loops

grantnet
```

Teach
```{r}
teachnet<- graph_from_adjacency_matrix(teach,
                                              mode="undirected", #Symmetrized
                                              weighted = NULL, #no weights
                                              diag = FALSE) #do not want loops

teachnet
```
Present
```{r}
presentnet<- graph_from_adjacency_matrix(present,
                                              mode="undirected", #Symmetrized
                                              weighted = NULL, #no weights
                                              diag = FALSE) #do not want loops

presentnet
```


# Add attributes

Now we want to add some attributes to our graphs. Since we have quite a few and I don't want to have to repeat the process for each, I'm going to attempt to save many lines of code and use functions.

Add all of our new igraph objects to a list.
```{r}
#From Bobbi's code:
NamedList<- function(...){
  names<- as.list(substitute(list(...)))[-1L]
  result<- list(...)
  names(result) <- names
  result
}


net_list<- NamedList(contactnet, anycollnet, researchnet, manuscriptnet, grantnet, teachnet, presentnet)
```


Function to add attributes.
```{r}
#from Bobbi's code:
NodeAtts<- function(x){
  add1<- set.vertex.attribute(x, "full_name", value=as.numeric(atts$full_name))
  add2<- set.vertex.attribute(add1, "position_color", value=atts$position_color)
  add3<- set.vertex.attribute(add2, "position", value=atts$position)
  add4<- set.vertex.attribute(add3, "location_color", value=atts$location_color)
  add5<- set.vertex.attribute(add4, "location", value=atts$location)
  add6<- set.vertex.attribute(add5, "connect_ease_color", value=atts$connect_rating_bi_color)
  add7<- set.vertex.attribute(add6, "connect_ease", value=atts$connect_rating_bi)
  add8<- set.vertex.attribute(add7, "referred_color", value=atts$referred_color)
  add9<- set.vertex.attribute(add8, "referred", value=atts$referred)
}
```

```{r}
net_list<- lapply(net_list, NodeAtts)
```

Here's an example now that we've added node attributes.
```{r}

summary(net_list$contactnet)
```

Now we save the full net list package as .rdata so we can access for analysis and visualization.
I'm making sure to package it with the year so we can differentiate.
```{r}
saveRDS(net_list, "net_list.rds")
```

