---
title: 'SNA Example Visualizations Template- Fellow'
author: "Rebekah Jacob"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

```{r}
library(tidyverse) #for data mgmt
library(DT) #for interactive html tables
library(igraph) #for network graph figures
library(arsenal) #for tableby function in member tables
```

```{r}
net_list<- readRDS("net_list.rds") #network list 

atts<- read_csv("Contacts/attribute-file.csv") #attributes
```

# Network Members

The table below describes members of the full network.
```{r results='asis'}

# controls for all Tableby tables throughout
mycontrols  <- tableby.control(total=FALSE, #don't want total column
                               test=FALSE, #don't want pvalue
                               numeric.stats = c("meansd"), digits=2L, cat.stats = c("countpct"))

person_char <- tableby(~
                  position +
                  location +
                  referred +
                  connect_rating +
                  connect_rating_bi, data=atts, control=mycontrols)

summary(person_char)
```

# Network Visualtization
## By Position

For network visualization, plots feature nodes that are sized by `degree` and color by `position`.

```{r}
#rescale for vertex size
rescale_fun<- function(nchar, low, high) {
  min_d<- min(nchar)
  max_d<- max(nchar)
  rscl<- ((high-low) * (nchar-min_d))/ (max_d-min_d)+low
  rscl
}


# with IDS
#plot/graph nodes by degree function
plot_deg_position_id <- function(x, y){
  position= c("Fellow", "Mentor")
color_pal= c("#15d4f7", "#f96e26")
  plot(x, edge.color="Light Gray", vertex.label=V(x)$full_name, vertex.label.cex=0.5, vertex.label.dist=1, vertex.color=V(x)$position_color, vertex.frame.color=NA, vertex.size=rescale_fun(degree(x), 5, 15), layout=layout_components, main=y)
  legend("bottomleft", legend=position, col=color_pal, bty="n", pch=20 , pt.cex = 3, cex = .8 , horiz = FALSE)
}

#Without IDs
plot_deg_position <- function(x, y){
    position= c("Fellow", "Mentor")
color_pal= c("#15d4f7", "#f96e26")
  plot(x, edge.color="Light Gray", vertex.label=NA, vertex.color=V(x)$position_color, vertex.frame.color=NA, vertex.size=rescale_fun(degree(x), 5, 15), layout=layout_components, main=y)
  legend("bottomleft", legend=position, col=color_pal, bty="n", pch=20 , pt.cex = 3, cex = .8 , horiz = FALSE)
}
```


### Contact
```{r}
png("Network figures/contact-by-position.png")
plot_deg_position(net_list$contactnet, "Contact by position")
dev.off()
plot_deg_position(net_list$contactnet, "Contact by position")
```

### Collaboration {.tabset .tabset-pills}
#### Research
```{r}
plot_deg_position(net_list$researchnet, "Research by position")
```

#### Manuscript
```{r}
plot_deg_position(net_list$manuscriptnet, "Manuscript by position")
```

#### Grant
```{r}
plot_deg_position(net_list$grantnet, "Grant by position")
```

#### Teach
```{r}
plot_deg_position(net_list$teachnet, "Teach by position")
```

#### Present
```{r}
plot_deg_position(net_list$presentnet, "Present by position")
```

#### Any Collaboration Combined
```{r}
png("Network figures/any-collab-by-position.png")
plot_deg_position(net_list$anycollnet, "Any Collaboration Combined by position")
dev.off()
plot_deg_position(net_list$anycollnet, "Any Collaboration Combined by position")
```



## By Location

For network visualization, plots feature nodes that are sized by `degree` and color by `location`.

```{r}
#rescale for vertex size
rescale_fun<- function(nchar, low, high) {
  min_d<- min(nchar)
  max_d<- max(nchar)
  rscl<- ((high-low) * (nchar-min_d))/ (max_d-min_d)+low
  rscl
}


# with IDS
#plot/graph nodes by degree function
plot_deg_location_id <- function(x, y){
  location= c("US", "Mex", "Canada")
color_pal= c("#f96e26", "#15d4f7", "#5e53f7")
  plot(x, edge.color="Light Gray", vertex.label=V(x)$full_name, vertex.label.cex=0.5, vertex.label.dist=1, vertex.color=V(x)$location_color, vertex.frame.color=NA, vertex.size=rescale_fun(degree(x), 5, 15), layout=layout_components, main=y)
  legend("bottomleft", legend=location, col=color_pal, bty="n", pch=20 , pt.cex = 3, cex = .8 , horiz = FALSE)
}

#Without IDs
plot_deg_location <- function(x, y){
  location= c("US", "Mex", "Canada")
color_pal= c("#f96e26", "#15d4f7", "#5e53f7")
  plot(x, edge.color="Light Gray", vertex.label=NA, vertex.color=V(x)$location_color, vertex.frame.color=NA, vertex.size=rescale_fun(degree(x), 5, 15), layout=layout_components, main=y)
  legend("bottomleft", legend=location, col=color_pal, bty="n", pch=20 , pt.cex = 3, cex = .8 , horiz = FALSE)
}
```

### Contact
```{r}
plot_deg_location(net_list$contactnet, "Contact by location")
```

### Collaboration {.tabset .tabset-pills}
#### Research
```{r}
plot_deg_location(net_list$researchnet, "Research by location")
```

#### Manuscript
```{r}
plot_deg_location(net_list$manuscriptnet, "Manuscript by location")
```

#### Grant
```{r}
plot_deg_location(net_list$grantnet, "Grant by location")
```

#### Teach
```{r}
plot_deg_location(net_list$teachnet, "Teach by location")
```

#### Present
```{r}
plot_deg_location(net_list$presentnet, "Present by location")
```

#### Any Collaboration Combined
```{r}
plot_deg_location(net_list$anycollnet, "Any Collaboration combined by location")
```
